\section{Introduction}\label{sec:introduction}

A major objective of data driven research on complex systems is the
identification of generic or universal statistical behavior. The tremendous
success of thermodynamics and statistical mechanics serves as an inspiration
when continuing this quest in complex systems beyond traditional physics.
Particularly interesting are large complex systems which consists of similar,
yet clearly distinguishable complex subsystems. The degree of universality
found in one particular subsystem can then be assessed if this type of
universality is also seen in another subsystem. Useful information on the impact
of specific system features on this universality may then be inferred.

Here, we carry out such a study for finance, because a tremendous amount of
data is available \cite{physicists_contribution}. Markets may be viewed as
macroscopic complex systems with an internal microscopic structure that is to a
large extent accessible by big data analysis \cite{complex_markets}.
Stock markets and foreign exchange markets are clearly distinct, but share many
common features. In previous analyses, we studied response functions in stock
markets to shed light on non-Markovian behavior. Here, we extend that to the
foreign exchange markets. To our surprise, we did not find such an
investigation in the literature. Hence, we believe that this study is a
rewarding effort. It helps to examine the behavior of the functions applied to
the foreign exchange market and it is suitable to compare the similarities and
differences to other markets.

The foreign exchange market is the most volatile, liquid and largest of all
financial markets
\cite{forex_liquidity,info_forex,forex_market_micro,intraday_forex}, and it has
a paramount importance for the world economy. It affects employment, inflation,
international capital flows, among others \cite{forex_structure}. The foreign
exchange market is a decentralized market without a common trading floor
\cite{info_forex,forex_market_micro,forex_structure,teach_spread}

The term pip (Price Increment Point) is commonly used in the foreign exchange
market instead of tick. The precise definition of a pip is a matter of
convention. Usually, it refers to the incremental value in the fifth non-zero
digit position from the left. It is not related to the position of the decimal
point. For example, one pip in the exchange rate USD/JPY of 124.21 would be
0.01, while one pip for EUR/USD of 1.1021 would be 0.0001
\cite{forex_market_micro,forex_structure,order_flow_forex,micro_eff}.

The foreign exchange market has attracted a lot of attention in the last 20
years. Electronic trading has changed an opaque market to a fairly transparent
one with transactions costs that are a fraction of their former level. The
large amount of data that is now available to the public make possible
different kinds of data analysis. Intense research is currently carried out in
different directions
\cite{forex_liquidity,info_forex,intraday_forex,forex_structure,teach_spread,forex_microstructure,electronic_forex,forex_algorithmic,curr_speculation,patterns_forex,eur_change_forex,spread_competition,political_forex,forex_volatility,local_forex,forex_inefficiency}.

McGroarty et al. \cite{micro_eff} found that smaller volumes cause larger
bid-ask spreads for technical reasons related to the measurement, whereas Hau
et al. \cite{eur_change_forex,eur_int_curr} claim that larger bid-ask spreads
caused smaller volumes due to the traders' behavior.

Burnside et al. \cite{curr_speculation} found the spreads to be between two and
four times larger for emerging market currencies than for developed country
currencies. According to Huang and Masulis \cite{spread_competition}, bid-ask
spreads increase when the foreign exchange market volatility increases, and
decrease when the competition between the dealers increases. Ding and Hiltrop
\cite{electronic_forex} showed that the Electronic Broking Services (EBS)
reduces spreads significantly, but dealers with information advantage tend to
quote relatively wider spreads. King \cite{spread_futures} analyzed the foreign
exchange futures market and observed that the number of transactions is
negatively related with bid-ask spread, whereas volatility in general is
positively related. Serbinenko and Rachev \cite{intraday_forex} focus on the
three major market characteristics, namely efficiency, liquidity and
volatility, and found that the market is efficient in a weak form. Menkhoff and
Schmeling \cite{local_forex} used orders from the Russian interbank for Russian
rouble/US dollar rate. They analyzed the price impact in different regions of
Russia, and found that regions that are centers of political and financial
decision making have high permanent price impact.

The price response functions measure price changes resulting from execution of
market orders. Specifically, they measure how a buy or sell order at time $t$
influences on average the price at a later time $t + \tau$. It was shown in
different works
\cite{components_spread_tokyo,dissecting_cross,r_walks_liquidity,subtle_nature,Bouchaud_2004,theory_market_impact,my_paper_response_financial,Wang_2016_avg,Wang_2016_cross}
that the price response functions increase to a maximum and then slowly
decrease as the time lag grows.

Little is known about price response functions in the foreign exchange markets:
\cite{forex_liquidity,forex_volatility,response_funct_fx}. Melvin and Melvin
\cite{forex_volatility} simulate their proposed model for different foreign
exchange markets region to analyze the impact of a one-standard-deviation shock
using impulse response functions. The general pattern of response was a fairly
steep drop over the first couple of days followed by a few days of gradual
decline until the response is not statistically different from zero. Mancini et
al. \cite{forex_liquidity} model the price impact and return reversal to
analyze liquidity. Their model predicts that more liquid assets should exhibit
narrower spreads and lower price impact.

To the best of our knowledge, no large-scale data analysis of response
functions for foreign exchange market has been carried out. Response functions
are important observables as they give information on non-Markovian behavior.
It is the purpose of the present study to close this gap. Based on a series of
detailed empirical results obtained on trade by trade data, we show that the
price response functions in the foreign exchange markets behave qualitatively
similar as the ones in correlated stocks markets. We consider different time
scales, years and currency pairs to compute the price response functions.
Finally, we shed light on the spread impact in the response functions for
foreign exchange pairs. We use a pip spread definition to group different
foreign exchange pairs and show that large pip spreads have stronger impact
on the response.To facilitate the reproduction of our results, the source code
for the data analysis is available in Ref. \cite{code}.

The paper is organized as follows: in Sect. \ref{sec:data_set} we present our
data set of foreign exchange pairs and briefly describe the physical and trade
time scale. We define the time scale to be used in Sect. \ref{sec:time_scale},
and compute the price response functions for the majors pairs in Sect.
\ref{sec:response_functions}. In Sect. \ref{sec:spread_impact} we show how
the spread impact the values of the response functions. Our conclusions follow
in Sect. \ref{sec:conclusion}.