\section{Data set}\label{sec:data_set}

In this study, we analyze foreign exchange pairs from the foreign exchange
market. In the foreign exchange market, the trading day begins in the markets
of Australia and Asia. Then the markets of Europe open and finishes late in the
afternoon in New York \cite{forex_structure,forex_market_micro}. The markets does not formally
closes during the week. Thus, using the New York time as reference, the market
opens on Sunday at 19h00 and closes on Friday at 17h00.

The foreign exchange financial data was obtained from
\href{www.histdata.com}{HistData.com}. We used a tick-by-tick database in
generic ASCII format for different years and currency pairs. The data comprises
the date time stamp (YYYYMMDD HHMMSSNNN), the best bid and best ask quotes
prices in the Eastern Standard Time (EST) time zone. No information about the
size of each transaction is provided. Also, the identity of the participants is
not given.

On physical time scale, for each exchange rate, we process the irregularly
spaced raw data to construct second-by-second price and volume series, each
containing 86,400 observations per day. For every second, the midpoint of best
bid and ask quotes or the transaction price of deals is used to construct
one-second log-returns.

To analyze the price response functions in Sect.
\ref{sec:response_functions}, we select the seven major currency pairs in
three different years (2008, 2014 and 2019). Table \ref{tab:majors} shows the
currency pairs with their corresponding symbols.

\begin{table}[htbp]
\centering
\begin{threeparttable}
\caption{Analyzed currency pairs.}
\begin{tabular*}{\columnwidth}{P{5cm}P{3cm}}
\toprule
\bf{Currency pair} & \bf{Symbol} \tabularnewline
\midrule
euro/U.S dollar& EUR/USD \tabularnewline
British pound/U.S dollar& GBP/USD \tabularnewline
Japanese yen/U.S dollar& JPY/USD \tabularnewline
Australian dollar/U.S dollar& AUD/USD \tabularnewline
U.S dollar/Swiss franc& USD/CHF \tabularnewline
U.S dollar/Canadian dollar& USD/CAD \tabularnewline
New Zealand dollar/U.S dollar& NZD/USD \tabularnewline
\bottomrule
\end{tabular*}
\label{tab:majors}
\end{threeparttable}
\end{table}

To analyze the spread impact in price response functions (Sect.
\ref{sec:spread_impact}), we select 46 currency pairs, in three different years
(2011, 2015 and 2019). The selected pairs are listed in Appendix
\ref{app:fx_pairs_spread}.

In order to avoid overnight effects and any artifact due to the opening and
closing of the foreign exchange market, we systematically discarded the first
ten and the last ten minutes of trading in a given week
\cite{Bouchaud_2004,large_prices_changes,spread_changes_affect,my_paper_response_financial,Wang_2016_cross}.
Therefore, we only consider trades of the same week from Sunday 19:10:00 to
Friday 16:50:00 New York local time. We will refer to this interval of time as
the ``market time".
