\section{Conclusion}\label{sec:conclusion}

Price response functions provide quantitative information on the deviation
from Markovian behavior. They measure price changes resulting from execution of
market orders. We used these functions in a large-scale data analysis for
foreign exchange markets. A study that to the best of our knowledge, was never
done before.

We analyzed price response functions in foreign exchange markets for different
years and different time scales. We used trade time scale and physical time
scale to compute the price response functions for the seven major foreign
exchange pairs for three different years. These major pairs are highly relevant
in the dynamics of the market.

The price response functions were analyzed according to the time scales. On
trade time scale, the signals were noisier. For both time scales we observe
that the signal for all the pairs increase to a maximum and then start to
slowly decrease. As the response functions can not grow indefinitely with the
time lag, they increase to a peak, to then decrease. However, for the year 2008
the shape of the signals is not well defined as the other years. The
increase-decrease behavior observed in the foreign exchange market was also
reported in correlated financial markets. These results show that the price
response functions conserve their behavior in different years and in different
markets.

In both scales, the more liquid pairs have a smaller price response function
compared with the non-liquid pairs. As the liquid pairs have more trades during
the market time, the impact of each trade is reduced. Comparing between years
and scales, the price response signal is stronger in past year than in recent
years. As algorithm trading has gain great relevance, in recent years the
quantity of trades has grown, and in consequence, the impact in the response
has decreased.

Finally, we checked the pip spread impact in price response functions for three
different years. We used 46 foreign exchange pairs and grouped them depending
on the conditions of the corresponding year analyzed. We employ the year
average pip spread of every pair for each year. For all the year and time
scales, the price response function signals were stronger for the groups of
pairs with larger pip spreads and weaker for the group of pairs with smaller
spreads. For the average of the price response functions, it was only possible
to see the increase-maximum-decrease behavior in the year 2015 in both scales,
and in the year 2019 in trade time scale. This suggest there is a lot of noise
in the cross and exotic pairs due to the lack of trading compared with the
majors. A general average price response behavior for each year and time scale
was spotted for the groups, suggesting a market effect on the foreign exchange
pairs in each year.